---
title: "p8105_hw3_az2587"
author: "Anyu Zhu"
date: "10/12/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(p8105.datasets)
library(tidyverse)
library(ggplot2)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
ggplot2.continuous.colour = "viridis",
ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1
```{r}
data("instacart")
instacart %>% 
  janitor::clean_names()

instacart %>% 
  distinct(aisle) 
```
There are **`r nrow(instacart)`** observations and **`r ncol(instacart)`** variables in the instacart dataset. The variables are `r colnames(instacart)`. 

Below are some specific questions related to the dataset：\\
How many aisles are there, and which aisles are the most items ordered from?
```{r}
aisle_ids = instacart %>% 
  distinct(aisle_id) 

aisle_best = instacart %>% 
  group_by(aisle_id) %>% 
  summarize(n_ordered = n()) %>% 
  mutate(aisle_rank = min_rank(desc(n_ordered))) %>% 
  filter(aisle_rank == 1)

aisle_best
```
There are in total **`r nrow(aisle_ids)`** aisles. The id of the aisle where most items are ordered from is **`r pull(aisle_best, aisle_id)`**. \\

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>% 
  group_by(aisle, department) %>% 
  summarize(n_ordered = n()) %>% 
  filter(n_ordered > 10000) %>% 
  #mutate(aisle = factor(aisle),
  #       aisle = fct_reorder(aisle, n_ordered)) %>% 
  ggplot(aes(x = reorder(aisle,n_ordered), y = n_ordered, fill = department)) + geom_col() + 
  labs(
    title = "number of orders in each aisle",
    x = "aisles",
    y = "number of orders"
  ) + 
  theme_minimal() +
  theme(axis.text.y = element_text(hjust = 1), 
        axis.text = element_text(size = 6),
        legend.position = "bottom",
        legend.text = element_text(size = 6)) + 
  coord_flip()
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart %>% 
  filter(aisle == c("baking ingredients", "dog food care","packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_times = n()) %>% 
  mutate(product_rank = min_rank(desc(n_times))) %>% 
  filter(product_rank == 1) %>% 
  select(-product_rank) %>% 
  knitr::kable()
```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = round(mean(order_hour_of_day), digits = 2)) %>% 
  pivot_wider(names_from = order_dow, values_from = mean_hour) %>% 
  knitr::kable()
```

## Problem 2

Data preparation
```{r}
data("brfss_smart2010")
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  mutate(
    response_rank = recode(response, `Poor` = 0, `Fair` = 1, `Good` = 2, `Very good` = 3,
                           `Excellent` = 4),
    response = fct_reorder(response, response_rank))
```

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
states_2002 = brfss %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  distinct(locationdesc) %>% 
  summarize(n_location_obs = n()) %>% 
  filter(n_location_obs >= 7)

states_2002
```
The states that were observed at 7 or more locations are `r pull(states_2002, locationabbr)`. 
```{r}
states_2010 = brfss %>% 
  filter(year == 2010) %>% 
  group_by(locationabbr) %>% 
  distinct(locationdesc) %>% 
  summarize(n_location_obs = n()) %>% 
  filter(n_location_obs >= 7)

states_2010
```
The states that were observed at 7 or more locations are `r pull(states_2010, locationabbr)`. 


Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
```{r}
excellent_df = brfss %>% 
  filter(response == "Excellent") %>% 
  group_by(locationabbr, year) %>% 
  summarize(mean_data_value = mean(data_value, na.rm = TRUE))

excellent_df
```

Make a “spaghetti” plot of this average value over time within a state.
```{r}
excellent_df %>% 
  ggplot(aes(x = year, y = mean_data_value, color = locationabbr)) +
  geom_point(alpha = 0.3) +
  geom_line(aes(group = locationabbr), alpha = 0.5)
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss %>% 
  group_by(year, locationabbr) %>% 
  filter((year == 2006 | year == 2010),
         locationabbr == "NY")%>% 
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid(. ~ year)
```




